### 2012-05-21 docs/makefiles/makefiles.tex
more build-system stuff

The \maestro\ build system uses features of GNU make, which is typically the 
default on systems.  The main macros that define the build process are 
split across several files.  The 4 main files are:
\begin{itemize}
\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}:
  This setups the basic macros, includes the options for the selected
  compiler, builds the list of object and source files, and defines
  the compile and link command lines.

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}
  This implements any changes to the compiler names and library
  locations necessary to build a parallel executable with MPI.

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}
  This creates the various build targets and specifies the rules for
  building the object files, the list of dependencies, and some other
  lesser-used targets (tags for editors, documentation, etc.)

\item {\tt MAESTRO/GMaestro.mak}
  This is a \maestro-specific file that gathers all of the various
  modules that are used to build a typical \maestro\ application
  and integrates with the \boxlib\ build system.  Every \maestro\
  problem's makefile will include this file.

@@

The \boxlib\ build system separates the compiler-specific information
from the machine-specific information---this allows for reuse of the
compiler information.  The only machine-specific parts of the build system
are for the MPI library locations, contained in {\tt GMakeMPI.mak}.

The \maestro\ build system relies on the vpath functionality of {\tt
make}.  In a makefile, the {\tt vpath} variable holds search path used
to locate source files.  When a particular file is needed, the
directories listed in {\tt vpath} are searched until the file is
found.  The first instance of a file is used.  We exploit this feature
by always putting the build directory first in {\tt vpath} (this is
done in {\tt GMakerules.mak}).  This means that if a source file is
placed in the build directory, that copy will override any other
version in the source path.

@@

files created at build-time: {\tt probin.f90}, {\tt build\_info.f90}, and {\tt network.f90}

@@

\subsection{Problem-specific Files}

\subsection{Defining Network, EOS, and Conductivity Routines}

@@

>>>>

The \maestro\ build system uses features of GNU make, which is
typically the default on systems.  The \maestro\ executable is built
in the problem's directory (one of the directories under {\tt
SCIENCE/}, {TEST\_PROBLEMS}, or {UNIT\_TESTS}).  This directory will
contain a makefile, {\tt GNUmakefile}, that includes all the necessary
information to build the executable.

The main macros that define the build process are split across several
files.  The 4 main files are:
\begin{itemize}
\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}:

  This setups the basic macros, includes the options for the selected
  compiler, builds the list of object and source files, and defines
  the compile and link command lines.

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}:

  This implements any changes to the compiler names and library
  locations necessary to build a parallel executable with MPI.

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}:

  This creates the various build targets and specifies the rules for
  building the object files, the list of dependencies, and some other
  lesser-used targets (tags for editors, documentation, etc.)

\item {\tt MAESTRO/GMaestro.mak}:

  This is a \maestro-specific file that gathers all of the various
  modules that are used to build a typical \maestro\ application
  and integrates with the \boxlib\ build system.  Every \maestro\
  problem's {\tt GNUmakefile} will include this file.

@@

\maestro\ gets the location of the \boxlib\ library through the 
{\tt BOXLIB\_HOME} variable.  This should be set as an environment
variable in your shell start-up files (e.g.\ {\tt .bashrc} or {\tt
.cshrc}).

The \boxlib\ build system separates the compiler-specific information
from the machine-specific information---this allows for reuse of the
compiler information.  The only machine-specific parts of the build system
are for the MPI library locations, contained in {\tt GMakeMPI.mak}.
The compiler flags for the various compilers are listed in the
files in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.

\subsection{Finding Source Files}

The \boxlib\ build system relies on the vpath functionality of {\tt
make}.  In a makefile, the {\tt vpath} variable holds search path used
to locate source files.  When a particular file is needed, the
directories listed in {\tt vpath} are searched until the file is
found.  The first instance of a file is used.  We exploit this feature
by always putting the build directory first in {\tt vpath} (this is
done in {\tt GMakerules.mak}).  This means that if a source file is
placed in the build directory, that copy will override any other
version in the source path.

In \maestro, the {\tt vpath} variable is set using the macros defined
in {\tt GMaestro.mak}.  A user does not need to set this variable
explicitly.  Additional source locations are added in the manner
described below (see \S~\ref{sec:make:otherfiles}).

\subsection{Dependencies}

@@

\subsection{Files Created at Compile-time}

Several files are created at build-time:
\begin{itemize}
\item {\tt probin.f90}

\item {\tt build\_info.f90}

\item {\tt network.f90} (only for the {\tt general\_null} network)
\end{itemize}

@@

\subsection{Problem-specific Files}
\label{sec:make:otherfiles}

\subsection{Defining EOS, Network, and Conductivity Routines}

Each \maestro\ problem needs to define an equation of state, a
reaction network, and a routine to compute the conductivities (for
thermal diffusion).  This is true even if the problem is not doing
reactions of thermal diffusion.  These definitions are specified
in the problem's {\tt GNUmakefile}.

\begin{itemize}
\item {\tt EOS\_DIR}:

  This variable points to the path (relative to {\tt fParallel/}) of
  the equation of state used for the build.  Choices that work
  with \maestro\ are:
  \begin{itemize}
  \item {\tt extern/EOS/helmeos}
  \item {\tt extern/EOS/gamma\_law\_general}
  \end{itemize}

\item {\tt CONDUCTIVITY\_DIR}:

  This variable points to the conductivity routine used for the build
  (relative to {\tt fParallel/}).  Choices that work with \maestro\
  are:
  \begin{itemize}
  \item {\tt extern/conductivity/constant}
  \item {\tt extern/conductivity/timmes\_stellar}
  \end{itemize}
  If diffusion is not being used for the problem, this should be set
  to {\tt extern/conductivity/constant}.

\item {\tt NETWORK\_DIR}:

  This variable points to the reaction network used for the build
  (relative to {\tt fParallel/}).  A lot of options are present in
  {\tt extern/networks/}.  A network is required even if you are not
  doing reactions, since the network defines the species that are
  advected and interpreted by the equation of state.

  A special choice, {\tt extern/networks/general\_null} is a general
  network that simply defines the properties of one or more species.
  This requires an inputs file, specified by {\tt
  GENERAL\_NET\_INPUTS}.  This inputs file is read at compile-time and
  used to build the {\tt network.f90} file that is compiled into the
  source.

@@

To see the contents of any variable in the build system, you can build
the special target {\tt print-{\em varname}}, where {\tt {\em
varname}} is the name of the variable.  For example, to see what the
Fortran compiler flags are, you would do:
\begin{verbatim}
make print-FFLAGS
\end{verbatim}
This would give (for {\tt gfortran}, for example):
\begin{verbatim}
FFLAGS is -Jt/Linux.gfortran/m -I t/Linux.gfortran/m -O2 -fno-range-check
\end{verbatim}
This functionality is useful for debugging the makefiles.
